---
title: "STAT598zHW1"
author: "Huiting Su"
date: "January 24, 2018"
output: html_document
---
## 1 Problem 1: The seq() function
1. The starting point is "from", and "to" is the end. Increment is "by", and "length.out" is the desired length.
```{r}
x <- seq(from=1, to=10, by=2)
x
seq(10, 1, length.out=5)
seq(10, 1, along.with=x)
```

2. If "to" is not specified, then the sequence length is  1:from. If "by" is not specified, the defaut value is 1. 
```{r}
seq(10)
seq(3,10)
```

3. When you pass seq() a vector, it returns a sequence of same length, starting from 1 and increment by 1.  
```{r}
seq(along.with=x)
```

4.  If the vector is of length 1, it is interpreted as 1:from, instead of "along.with".
```{r}
y <- seq.int(3,3,1)
y
seq(y)
```

5.It is preferable to use seq_along() in this case.
```{r}
seq_along(x)
seq_along(y)
```

6.
```{r}
z <- as.integer(seq_len(8)+2)
z
typeof(z)
```
7. For positive case it is 1:num. For negative case it is error.
```{r}
seq_len(8)
#seq_len(-8)
```

## 2 Problem 2: Vectors
1. 
```{r}
letters
summary(letters)
```
2. 
```{r}
typeof(letters)
```

3. 
```{r}
len <- length(letters)
len
```

4. 
```{r}
alterletter <- letters[seq(1,len,2)]
alterletter
```

5. 
```{r}
backletter <- letters[seq(len,1)]
backletter
```
6. 
```{r}
alterback <- letters[seq(len,1,-2)]
alterback
```

7. 
```{r}
matrix(letters[1:16],4,4)
```

8. Built-in Constants are LETTERS, letters, month.abb, month.name, pi.
```{r}
c(typeof(month.abb),typeof(month.name),typeof(pi))
```

## 3 Problem 3: Matrices
1. 
```{r}
m <- matrix(runif(20),4,5)
```

2. It is an array, and it is not an 1 dimension matrix. There is no dim attribute.
```{r}
m[3,]
dim(m[3,])
```

3. In this way, the result is not coerced to the lowest possible dimension, which means it is still a matrix.
```{r}
m[3,,drop=FALSE]
dim(m[3,,drop=FALSE])
```

4. 
```{r}
sum(m)
```

5.
```{r}
m <- m/sum(m)
sum(m)
```

6.
```{r}
rowSums(m)
colSums(m)
```

7.
```{r}
m <- m/rowSums(m)
rowSums(m)
```

8.Transpose.
```{r}
dim(t(m))
```

9.
```{r}
m <- t(t(m)/colSums(m))
colSums(m)
```

10. First every other element mving along the columns, then rows.
```{r}
m2 <- matrix(runif(2*9),2,9)
m2
m2[seq(1,18,2)]
t(m2)[seq(1,18,2)]
```

11. A vector of "1" and vector 2 are column bind to be a matrix. This matrix is then column bind with vector 3. However, by matrix indexing, vector 3 is put in between vector 1 and vector 2.  
```{r}
m <- cbind(1, 1:7) 
# the ’1’ (= shorter vector) is recycled
m
m <- cbind(m, 8:14)[, c(1, 3, 2)] 
# insert a column
m

```




