---
title: "Stat598 HW3 -- dataframes and ggplot2"
author: "Huiting Su"
date: "February 19, 2018"
output: html_document
---
(a) Install
```{r, warning=FALSE, message=FALSE}
#install.packages('maps')
library(maps)
library(ggplot2)
library(dplyr)
#install.packages('ggforce')
library(ggforce)
```
(b) State map dataframe.
```{r}
states_map <- map_data(map = "state")
dim(states_map)
colnames(states_map)
```
(c) Unique value.
```{r}
unique(states_map$region)
```

(d)
```{r}
g <- ggplot(states_map, aes(x=long, y=lat, color=region)) + geom_point() + guides(colour=FALSE)
g
```

(e) Connect the points with lines.
```{r}
g + geom_line()
g + geom_path()
```

We can see geom_path is more appropriate. According to the help doc, geom_path() connects the observations in the order in which they appear in the data. geom_line() connects them in order of the variable on the x axis.

(f)
```{r}
state2 <- states_map[states_map$region=="california" | states_map$region== "colorado",]
ggplot(state2, aes(x=long, y=lat, color=region)) + geom_point()  + geom_path()
```
(g)
```{r}
set.seed(4365)
state2_random <- state2[sample(1:dim(state2)[1]),]
ggplot(state2_random, aes(x=long, y=lat, color=region)) + geom_point()  + geom_path()
```

(h)
```{r}
ggplot() + 
    geom_map(map = states_map, map_id=states_map$region, data = states_map, aes(fill=group))+ 
    expand_limits(x = states_map$long, y = states_map$lat)
```

(i) Add State ID.
```{r}
regions<- unique(states_map$region)
n <- length(regions)
regions <- cbind.data.frame(stateid=1:n, region=regions)
states_map <- merge(states_map, regions)
```

```{r}
ggplot() + 
    geom_map(map = states_map, map_id=states_map$region, data = states_map, aes(fill=stateid))+ 
    expand_limits(x = states_map$long, y = states_map$lat)
```

(j) Add murder arrests per 100,000 people for different states.
```{r}
data(state)
states_map$InfoValue <- state.x77[match(states_map$region,tolower(rownames(state.x77))),5]
```

(k) Label and bind dataframes.
```{r}
states_map$InfoType = "Murder"
states_map2 <- states_map
states_map2$InfoType <- "Grad"
states_map2$InfoValue <- state.x77[match(states_map$region,tolower(rownames(state.x77))),6]
statesbind <- rbind(states_map, states_map2)
```


(l) Plot.
```{r}
#ggplot() + 
#    geom_map(map = statesbind, map_id=statesbind$region, data = statesbind, aes(fill=InfoValue))+ 
#    expand_limits(x = statesbind$long, y = statesbind$lat) + facet_grid(.~InfoType)
```
Unfortunately, the method mentioned in the HW instruction does not work for me, so I use "melt" function instead.
```{r}
Infos <- data.frame(state = tolower(rownames(state.x77)), state.x77[,5:6])
Infosmelt <- reshape2::melt(Infos, id = 1)
ggplot(Infosmelt, aes(map_id = state)) +
    geom_map(aes(fill = value), map = states_map) +
    expand_limits(x = states_map$long, y = states_map$lat) +
    facet_grid( ~ variable)
```


(m) Calculate dataframe of circle center.
```{r}
by_state <- group_by(states_map, State= region)
states_circle <- summarize(by_state, MeanLong = mean(long), MeanLat = mean(lat))
states_circle$ArrentCount <- state.x77[match(states_circle$State,tolower(rownames(state.x77))),3]
```

```{r, warning=FALSE, message=FALSE}
ggplot() + 
    geom_circle(aes(x0=MeanLong, y0=MeanLat, r=ArrentCount, fill=State), data=states_circle) + 
    guides(fill=FALSE)
```

(n)
```{r}
X <- unique(states_map$region)
fun <- function(statename) 
{
    MeanLong <- mean(states_map$long[states_map$region==statename])
    MeanLat <- mean(states_map$lat[states_map$region==statename])
    c(MeanLong, MeanLat) 
}
circle2 <- data.frame(t(sapply(X, fun)))
colnames(circle2) <- c('MeanLong', 'MeanLat')
head(circle2)
head(states_circle)
```

The two dataframes are the same.




