---
title: "K-nearest Neighbours"
author: "Huiting Su"
date: "March 6, 2018"
output: html_document
---
This is a practice of k-nearest neighbours.  

####(a) K-nearst Neighbours skeleton:
x' <- x value of new test datapoint  
train <- training data (X,Y)  
distance <- train[order((train$X-x'))]   #calculate distance and sort  
cfreq <- count(distance[1:k], vars= 'Y')  #count frequency of each label  
y' <- cfreq[cfreq\$freq == max]\$Y  #the most frequent label is our prediction  

####(b) Digits dataset contains 1100 examples of each digit.
```{r, warning=FALSE, message=FALSE}
#install.packages(RnavGraphImageData)
library(RnavGraphImageData)
data(digits)
library(plyr)
library(magrittr)
```

####(c) Visualize digits.
```{r}
plot_digit <- function(x){
    m <- matrix(x, nrow=16, ncol=16, byrow = TRUE)
    m <- apply(m, 1, rev)
    m <- t(m)
    image(m, col = gray(0:255/255))
}
```

####(d) Plot first instance of each digit.
```{r, fig.width=5, fig.height=5, message=FALSE}
apply(digits[, seq(1, 5500, 1100)], 2, plot_digit)
plot_digit(digits[, 5502])
plot_digit(digits[, 6603])
apply(digits[, seq(7701, 11000, 1100)], 2, plot_digit)
```

####(e) Convert to smaller dataset.
```{r}
get_digits <- function(select_digs, size){
    index <- numeric()
    select_digs[select_digs==0] <- 10
    start <- (select_digs-1)*1100+1
    for(ii in 1:length(start)){
        index <- c(index, seq(start[ii], length.out = size))
    }
    return(digits[,index])
}
```

```{r}
select_digs <- c(0,8)
size <- 100
my_train <- get_digits(select_digs, size)
```

####(f) Calculate Euclidean distance.
```{r}
euc_dist <- function(a, b) (a-b)^2 %>% sum %>% sqrt
```

####(g) Calculate distance between new digit and every element in my_train.
```{r}
new_digit <- my_train[1]
distances <- laply(my_train, euc_dist, new_digit)
head(distances)
```

####(h) Function to find k-nearest neighbours.
```{r}
get_knn <- function(k, my_train, new_digit){
    distances <- laply(my_train, euc_dist, new_digit)
    indices <- sort(distances, index.return = TRUE)$ix
    return(indices[1:k])
}
```

####(i) Get label.
```{r}
my_labels <- rep(c(0, 8), each= 100)

get_label <- function(kindices){
    labels <- my_labels[kindices]
    freq <- table(labels)
    return(names(which.max(freq)))
}
```

####(j) Wrap previous functions.
```{r}
my_knn <- function(k, my_train, my_labels, test_ip){
    kindices <- get_knn(k, my_train, test_ip)
    return(get_label(kindices))
}
```

####(k) Test the function on training set.
```{r, cache=TRUE}
k <- 5
prediction <- laply(my_train, my_knn, my_train = my_train, k = k, my_labels = my_labels)
paste("my_knn got wrong ", sum(prediction[1:100] != my_labels[1:100]), " 0's.")
paste("my_knn got wrong ", sum(prediction[101:200] != my_labels[101:200]), " 8's.")
```

####(l) Test the function on a testing set.
```{r, cache=TRUE}
my_test <- digits[,c(10000:10099, 7900:7999)]
test_labels <- rep(c(0, 8), each =100) 
prediction <- laply(my_test, my_knn, my_train = my_train, k = k, my_labels = my_labels)
paste("my_knn got wrong ", sum(prediction[1:100] != test_labels[1:100]), " 0's.")
paste("my_knn got wrong ", sum(prediction[101:200] != test_labels[101:200]), " 8's.")
```

Apply my_knn to one hundred 5's.
```{r, cache=TRUE}
my_test5 <- digits[,4900:4999]
prediction <- laply(my_test5, my_knn, my_train = my_train, k = k, my_labels = my_labels)
table(prediction)
```

























