---
title: "Reactive Programming and Monte Carlo Methods"
author: "Huiting Su"
date: "April 29, 2018"
output: html_document
---
```{r, message=FALSE, warning=FALSE}
library(shiny)
library(ggvis)
```

## Problem 1: The law of large numbers

### (a) Show evolution of the running mean with time.

```{r}
dat <- data.frame(indx = 0, value = c(0), running_mean = c(0))
ddat <- reactive({
 invalidateLater(200, NULL);
 len <- length(dat$indx) + 1;
 dat[len,] <<- c(len, rnorm(1),0)
 dat$running_mean[len] <<- mean(dat$value)
 dat
})
ddat %>% ggvis(x = ~indx, y = ~running_mean, key := ~indx) %>%
 layer_paths()
```

### (b) Show histogram of value.
```{r, message = FALSE}
ddat %>% ggvis(x = ~value) %>%
 layer_histograms()
```

## Problem 2: Monte Carlo Sampling
### (a) Estimate distance between two points.
```{r}
set.seed(876523)
p1 <- matrix(rnorm(5000*2, mean=0), 5000, 2)
p2 <- matrix(rnorm(5000*2, mean=1), 5000, 2)
distance <- sqrt((p2[,1]-p1[,1])^2 + (p2[,2]-p1[,2])^2)
avg_dis <- mean(distance)
print(avg_dis)
```

### (b) Repeat 1000 times, and show histogram of the estimations.
```{r}
estimate_dis <- function(rep){
    p1 <- matrix(rnorm(5000*2, mean=0), 5000, 2)
    p2 <- matrix(rnorm(5000*2, mean=1), 5000, 2)
    distance <- sqrt((p2[,1]-p1[,1])^2 + (p2[,2]-p1[,2])^2)
    avg_dis <- mean(distance)
    result[rep] <<- avg_dis
}

result <<- numeric()
set.seed(73487)
for(i in 1:1000){
    estimate_dis(i)
}
result <- as.data.frame(result)

ggvis(result) %>%
 layer_histograms(x=~result, fill.hover:='red')
```

## Problem 3: Importance Sampling

### (a) Mean and standard deviation of the sum of 100 fair dice.

$\mu_x = 100 \cdot 3.5 = 350$
$\sigma^2_x = 100 \cdot 2.917 = 291.7$ 
$\sigma_x = \sqrt{291.7} = 17.08$ 

### (b) Show the simulation result match (a)

```{r}
set.seed(73487)
dice <- rmultinom(1000, 100, c(1,1,1,1,1,1)/6)
dice_sum <- t(dice) %*% c(1,2,3,4,5,6)

avg_dice <- mean(dice_sum)
sd_dice <- sd(dice_sum)
cat(avg_dice, sd_dice)

dice_sum <- as.data.frame(dice_sum)
ggvis(dice_sum) %>%
 layer_histograms(x=~V1, fill.hover:='red')

sum(dice_sum > 450)
```

### (c) Calculate probability.
```{r}
pnorm(450, mean = avg_dice, sd = sd_dice, lower.tail = FALSE, log.p = TRUE)
```

### (d) Compare with simulation probability.

```{r}
sum(dice_sum > 450)/1000
```

### (e) Simulate biased dice.
```{r}
set.seed(73487)
die_biased <- rmultinom(1000, 100, c(1,2,3,4,5,6)/21)
biased_sum <- die_biased[1,] + 2 * die_biased[2,] + 3 * die_biased[3,] + 4 * die_biased[4,] + 5 * die_biased[5,] + 6 * die_biased[6,]

avg_biased <- mean(biased_sum)
sd_biased <- sd(biased_sum)
cat(avg_biased, sd_biased)

biased_sum <- as.data.frame(biased_sum)
ggvis(biased_sum) %>%
 layer_histograms(x=~biased_sum, fill.hover:='red')

sum(biased_sum > 450)
```

### (f) Probability.

```{r}
logp =  log(1/21)*die_biased[1,] + log(2/21)*die_biased[2,]  + log(3/21)*die_biased[3,] + log(4/21)*die_biased[4,] + log(5/21)*die_biased[5,] + log(6/21)*die_biased[6,]
head(logp)
```

### (g) Log-probability under fair dice.
```{r}
logpfair =  log(1/6)*colSums(dice)
head(logpfair)
```

### (h)
```{r}
w = logpfair / logp
indicator = biased_sum > 450
sum(indicator * w)/1000
```




